### 2017-02-13 robust-software.tex
Tidying up the formatting a bit

%% Include all macros below
\newcommand{\fixme}[2]{\textsc{\textbf{{#1}: {#2}}}}
\newcommand{\recommend}[1]{\textit{#1}}
\newcommand{\withurl}[2]{{#1}\footnote{\texttt{#2}}}
\newcommand{\rulemajor}[1]{\section{#1}}
\newcommand{\ruleminor}[1]{\textbf{#1}}

\begin{document}
\vspace*{0.2in}

@@

\linenumbers

@@

Being unable to
replicate results is so common that one publication refers to it as ``a rite of
passage''~\cite{baker2016}.
The root cause of this problem is that most research software
is essentially a prototype, and therefore is not
\emph{robust}. The lack of
robustness in published, distributed software
leads to duplicated efforts with little practical benefit,
which slows the pace of research~\cite{prabhu2011,lawlor2015}.
Not all software needs to be robust enough to run for other people,
but software that is published and distributed to the community should be.
Bioinformatics software repositories \cite{ison2016,brazas2012} catalogue dozens to
hundreds of tools that perform similar tasks:
for example,
in 2016 the Bioinformatics Links Directory included 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.
Some of these tools are legitimate efforts to improve the state-of-the-art, but
often they are difficult to install and run~\cite{stajich2002,Seemann2013}, and are effectively abandoned
after publication~\cite{nekrutenko2012}.

@@

\item \textit{List all input and output files},
even those considered self-explanatory.
Link to specifications for standard formats
and list the required fields and acceptable values in other files.
If there is no rigorous format, explain what the sections mean.
These files are often full of valuable information that can be
mined for the user's specific purpose.

@@

In order to save typing, it is often convenient to allow users to
specify an input or output \emph{directory}, and then require that there
be files with particular names in that directory. This practice, which
is sometimes called ``convention over configuration'', is used by many
software frameworks, such as WordPress and Ruby on Rails, and often
strikes a good balance between adaptability and consistency.

@@

Every package should come with a small test script for users to run
after installation. Its purpose is not only to check that the software
is working correctly (although that is extremely helpful), but also to
ensure that it works at all. This test script can also serve as a
working example of how to run the software.

In order to be useful, \ruleminor{make the test script easy to find and run}.
Many build systems will also run unit tests if provided them at compile time.
If the build system is not amenable to testing, a
single file in the project's root directory named \texttt{runtests.sh}
or something equally obvious is a 
good solution that lets new users build their analysis from a working script.
For example, with its distribution, 
HISAT2 includes a full set of very small files,
and a 'Getting Started with HISAT2' section in its manual 
that leads you through the entire data lifecycle~\cite{pertea2016}.

@@

The usage message tells users what the program could do.
It is equally important for the program to tell users what it actually did.
Accordingly,
when the program starts, it should \ruleminor{echo all parameters and software
versions to standard out or a log file alongside the results}. This
feature supports greater reproducibility because any result can be
replicated with only the previous output files as reference.

@@

\bibliography{robust-software}

>>>>

%% Include all macros below
\newcommand{\fixme}[2]{\textsc{\textbf{{#1}: {#2}}}}
\newcommand{\recommend}[1]{\textit{#1}}
\newcommand{\withurl}[2]{{#1}\footnote{\texttt{#2}}}
\newcommand{\rulemajor}[1]{\section{#1}}
\newcommand{\ruleminor}[1]{\textbf{#1}}
\begin{document}
\vspace*{0.2in}

@@

% \linenumbers

@@

Being unable to
replicate results is so common that one publication refers to it as ``a rite of
passage''~\cite{baker2016}.
The root cause of this problem is that most research software
is essentially a prototype, and therefore is not
\emph{robust}. The lack of
robustness in published, distributed software
leads to duplicated efforts with little practical benefit,
which slows the pace of research~\cite{prabhu2011,lawlor2015}.
Bioinformatics software repositories \cite{ison2016,brazas2012} catalogue dozens to
hundreds of tools that perform similar tasks:
for example,
in 2016 the Bioinformatics Links Directory included 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.
Some of these tools are legitimate efforts to improve the state-of-the-art, but
often they are difficult to install and run~\cite{stajich2002,Seemann2013}, and are effectively abandoned
after publication~\cite{nekrutenko2012}.

@@

\item \textit{List all input and output files},
even those considered self-explanatory.
Link to specifications for standard formats
and list the required fields and acceptable values in other files.
If there is no rigorous definition for a format, explain its parts
as clearly as possible in plain English.

@@

In order to save typing, it is often convenient to allow users to
specify an input or output \emph{directory}, and then require that there
be files with particular names in that directory. This practice is an
example of ``convention over configuration'', a principle is used by
software frameworks such as WordPress and Ruby on Rails that often
strikes a good balance between adaptability and consistency.

@@

Every package should come with a set of tests for users to run
after installation. Its purpose is not only to check that the software
is working correctly (although that is extremely helpful), but also to
ensure that it works at all. This test script can also serve as a
working example of how to run the software.

In order to be useful, \ruleminor{make the tests easy to find and run}.
Many build systems will also run unit tests if provided them at compile time.
For users, or
if the build system is not amenable to testing, 
provide a working script
in the project's root directory named \texttt{runtests.sh}
or something equally obvious. 
This lets new users build their analysis from a working script.
For example, with its distribution, 
HISAT2 includes a full set of very small files,
and a 'Getting Started with HISAT2' section in its manual 
that leads you through the entire data lifecycle~\cite{pertea2016}.

@@

The usage message tells users what the program could do.
It is equally important for the program to tell users what it actually did.
Accordingly,
when the program starts, it should \ruleminor{echo all parameters and software
versions to standard out or a log file alongside the results} to increase
the reproducibility of that step.

@@

\begin{thebibliography}{10}

\bibitem{prins2015}
Prins P, de~Ligt J, Tarasov A, Jansen RC, Cuppen E, Bourne PE.
\newblock Toward effective software solutions for big biology.
\newblock Nature Biotechnology. 2015;33(7):686--687.
\newblock doi:{10.1038/nbt.3240}.

\bibitem{baker2016}
Baker M.
\newblock 1,500 scientists lift the lid on reproducibility.
\newblock Nature. 2016;533(7604):452--454.
\newblock doi:{10.1038/533452a}.

\bibitem{prabhu2011}
Prabhu P, Jablin TB, Raman A, Zhang Y, Huang J, Kim H, et~al.
\newblock A Survey of the Practice of Computational Science.
\newblock In: State of the Practice Reports. SC '11. New York, NY, USA: ACM;
  2011. p. 19:1--19:12.

\bibitem{lawlor2015}
Lawlor B, Walsh P.
\newblock Engineering bioinformatics: building reliability, performance and
  productivity into bioinformatics software.
\newblock Bioengineered. 2015;6(4).
\newblock doi:{10.1080/21655979.2015.1050162}.

\bibitem{ison2016}
Ison J, Rapacki K, Ménager H, Kalaš M, Rydza E, Chmura P, et~al.
\newblock Tools and data services registry: a community effort to document
  bioinformatics resources.
\newblock Nucleic Acids Research. 2016;44(D1):D38--D47.
\newblock doi:{10.1093/nar/gkv1116}.

\bibitem{brazas2012}
Brazas MD, Yim D, Yeung W, Ouellette BFF.
\newblock A decade of web server updates at the bioinformatics links directory:
  2003-–2012.
\newblock Nucleic Acids Research. 2012;doi:{10.1093/nar/gks632}.

\bibitem{stajich2002}
Stajich JE, Block D, Boulez K, Brenner SE, Chervitz SA, Dagdigian C, et~al.
\newblock The Bioperl Toolkit: Perl Modules for the Life Sciences.
\newblock Genome Research. 2002;12(10):1611--1618.
\newblock doi:{10.1101/gr.361602}.

\bibitem{Seemann2013}
Seemann T.
\newblock Ten recommendations for creating usable bioinformatics command line
  software.
\newblock GigaScience. 2013;2(1):15.
\newblock doi:{10.1186/2047-217X-2-15}.

\bibitem{nekrutenko2012}
Nekrutenko A, Taylor J.
\newblock Next-generation sequencing data interpretation: enhancing
  reproducibility and accessibility.
\newblock Nature ReviewsGenetics. 2012;13(9):667--72.
\newblock doi:{10.1038/nrg3305}.

\bibitem{atwood2015}
Atwood TK, Bongcam-Rudloff E, Brazas ME, Corpas M, Gaudet P, Lewitter F, et~al.
\newblock GOBLET: The Global Organisation for Bioinformatics Learning,
  Education and Training.
\newblock PLoS Computational Biology. 2015;doi:{10.1371/journal.pcbi.1004143}.

\bibitem{sanders2008}
Sanders R, Kelly D.
\newblock Dealing with Risk in Scientific Software Development.
\newblock Software, IEEE. 2008;25(4):21--28.
\newblock doi:{10.1109/ms.2008.84}.

\bibitem{afgan2016}
Afgan E, Baker D, van~den Beek M, Blankenberg D, Bouvier D, Čech M, et~al.
\newblock The Galaxy platform for accessible, reproducible and collaborative
  biomedical analyses: 2016 update.
\newblock Nucleic Acids Research. 2016;44(W1):W3--W10.
\newblock doi:{10.1093/nar/gkw343}.

\bibitem{howe2012}
Howe B.
\newblock Virtual Appliances, Cloud Computing, and Reproducible Research.
\newblock Computing in Science Engineering. 2012;14(4):36--41.
\newblock doi:{10.1109/MCSE.2012.62}.

\bibitem{gentleman2004}
Gentleman RC, Carey VJ, Bates DM, Bolstad B, Dettling M, Dudoit S, et~al.
\newblock Bioconductor: open software development for computational biology and
  bioinformatics.
\newblock Genome Biology. 2004;5(10):1--16.
\newblock doi:{10.1186/gb-2004-5-10-r80}.

\bibitem{varoquaux2015}
Varoquaux G. Software for reproducible science: let’s not have a
  misunderstanding; 2015.
\newblock
  \url{http://gael-varoquaux.info/programming/software-for-reproducible-science-lets-not-have-a-misunderstanding.html}.

\bibitem{brown2013}
Brown CT. Replication, reproduction, and remixing in research software; 2013.
\newblock \url{http://ivory.idyll.org/blog/research-software-reuse.html}.

\bibitem{wilson2014}
Wilson G, Aruliah DA, Brown CT, Hong NPC, Davis M, Guy RT, et~al.
\newblock Best Practices for Scientific Computing.
\newblock PLoS Biology. 2014;12(1):e1001745.
\newblock doi:{10.1371/journal.pbio.1001745}.

\bibitem{wilson2016}
Wilson G, Bryan J, Cranston K, Kitzes J, Nederbragt L, Teal TK.
\newblock Good Enough Practices in Scientific Computing.
\newblock arxivorg. 2016;abs/1609.00037.

\bibitem{noble2009}
Noble WS.
\newblock {A Quick Guide to Organizing Computational Biology Projects}.
\newblock PLoS Computational Biology. 2009;5(7).
\newblock doi:{10.1371/journal.pcbi.1000424}.

\bibitem{blischak2016}
Blischak JD, Davenport ER, Wilson G.
\newblock A Quick Introduction to Version Control with Git and GitHub.
\newblock PLOS Computational Biology. 2016;12(1):1--18.
\newblock doi:{10.1371/journal.pcbi.1004668}.

\bibitem{karimzadeh2016}
Karimzadeh M, Hoffman MM. Creating great documentation for bioinformatics
  software; 2016.
\newblock \url{http://hdl.handle.net/1807/73111}.

\bibitem{Johnson1997}
Johnson M.
\newblock Building a Better ReadMe.
\newblock Technical Communication. 1997;44(1):28--36.

\bibitem{gnustandards}
{\relax Free Software Foundation}. GNU Coding Standards; 2016.
\newblock \url{https://www.gnu.org/prep/standards/standards.html}.

\bibitem{posix2016}
{\relax The IEEE}, {\relax The Open Group}. The Open Group Base Specifications
  Issue 7 IEEE Std 1003.1-2008. 12. Utility Conventions; 2016.
\newblock
  \url{http://pubs.opengroup.org/onlinepubs/9699919799/basedefs/V1_chap12.html}.

\bibitem{pertea2016}
Pertea M, Kim D, Pertea GM, Leek JT, Salzberg SL.
\newblock Transcript-level expression analysis of RNA-seq experiments with
  HISAT, StringTie and Ballgown.
\newblock Nature Protocols. 2016;11(9):1650--1667.

\bibitem{steinmacher2015}
Steinmacher I, Silva MAG, Gerosa MA, Redmiles DF.
\newblock A systematic literature review on the barriers faced by newcomers to
  open source software projects.
\newblock Information and Software Technology. 2015;59:67 -- 85.
\newblock doi:{http://dx.doi.org/10.1016/j.infsof.2014.11.001}.

\end{thebibliography}
